%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{report}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{graphicx}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{caption}
\doublespacing

\usepackage{geometry}
\geometry{margin=1.1in}




\let\oldnl\nl% Store \nl in \oldnl
\newcommand{\nonl}{\renewcommand{\nl}{\let\nl\oldnl}}% Remove line number for one line
\SetNlSty{textbf}{}{.}

\newcommand{\declareAlgo}[1]{\refstepcounter{algocf}\label{#1}} %

\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{problem}{Problem}
\newtheorem{theorem}{Theorem}
\newtheorem{question}{Question}
\newcommand{\commentOut}[1]{}


\makeatletter
\newcommand{\RemoveAlgoNumber}{\renewcommand{\fnum@algocf}{\AlCapSty{\AlCapFnt\algorithmcfname}}}
\newcommand{\RevertAlgoNumber}{\algocf@resetfnum}
\makeatother


\newcommand{\thesisTitle}{{\bf Data Structures and Algorithms for the\\
Identification of Biological Patterns}} 
\newcommand{\thesisAuthorName}{Marius Nicolae}

\title{\thesisTitle}


\author{\thesisAuthorName{}\\ marius.nicolae@engr.uconn.edu\\
Major Advisor: Prof. Sanguthevar Rajasekaran\\
Associate Advisors: Prof. Ion M\u{a}ndoiu and Prof. Yufeng Wu\\
Department of Computer Science and Engineering, University of Connecticut}

\begin{document}

\pagenumbering{gobble}



\thispagestyle{empty}
\begin{center}
\begin{minipage}{\linewidth}
    \centering
%Thesis title
    {\uppercase{\Large \thesisTitle \par}}
    %\vspace{1cm}
%Author's name
    {\Large \thesisAuthorName, PhD\par}
    %\vspace{1cm}

    {\Large University of Connecticut, 2016\par}
\end{minipage}
\end{center}
\vspace{1cm}


%location of RNA degradation signals, alternative splicing sites, etc.

%\begin{abstract}
This thesis studies the following problems:

{\bf 1. Planted Motif Search.} Discovering patterns in biological
sequences is a crucial process that has resulted in the determination of open
reading frames, gene promoter elements, intron/exon splicing sites, SH RNAs,
etc. We study the $(\ell,d)$ motif search problem or Planted Motif Search (PMS). 
PMS receives as input $n$ strings and two integers $\ell$ and $d$.
It returns all sequences $M$ of length $\ell$ that occur in each input string, 
where each occurrence differ from $M$ in at most $d$ positions.
Another formulation is quorum PMS (qPMS), where $M$ appears in at least 
$q\%$ of the strings. We developed qPMS9, an efficient parallel exact 
qPMS algorithm for DNA and protein datasets. 


{\bf 2. Suffix Array Construction.} The suffix array is a data
structure that finds numerous applications in string processing problems for both
linguistic texts and biological data. The suffix array consists of the sorted
suffixes of a string. There are several linear time suffix array construction algorithms known in the literature.
However, one of the fastest algorithms in practice has a worst case
run time of $O(n^2)$. We developed an efficient algorithm called RadixSA
that has a worst case run time of  $O(n\log{n})$ and is one of the fastest
algorithms to date. RadixSA introduces an idea that may find independent
applications as a speedup technique for other algorithms.

        
{\bf 3. Pattern Matching with Mismatches}. We consider several
variants of the pattern matching with mismatches problem. Given a text $T=t_1
t_2\cdots t_n$ and a pattern $P=p_1p_2\cdots p_m$, we investigate the following problems: 
1) {\em Pattern matching with mismatches:} for every alignment $i, 1\leq i \leq
n-m+1$ output the distance between $P$ and $t_i t_{i+1}\cdots t_{i+m-1}$, and 
2) {\em Pattern matching with $k$ mismatches:} output those alignments $i$
where the distance is at most $k$. The distance metric used is the Hamming
distance. Variants of these problems allow for wild cards in
the text or the pattern. For these problems we offer novel deterministic,
randomized and approximation algorithms.

Source code relevant to these results is available at
https://github.com/mariusmni/.
%\end{abstract}
\clearpage


\pagenumbering{Roman} 

%\thispagestyle{empty}
\begin{center}
\begin{minipage}{\linewidth}
    \centering
%Thesis title
    {\Large \thesisTitle \par}
    \vspace{1cm}
%Author's name
    {\Large \thesisAuthorName\par}
    \vspace{2cm}
    {\Large Dipl. Ing., Politehnica University of Bucharest, 2009\par}
    {\Large M.Sc., University of Connecticut, 2011\par}
    \vspace{5cm}

    {\Large A Dissertation\par}
    {\Large Submitted in Partial Fulfillment of the\par}
    {\Large Requirements for the Degree of\par}
    {\Large Doctor of Philosophy\par}
    {\Large at the\par}
    {\Large University of Connecticut\par}
    \vspace{1cm}
    {\Large 2016\par}
\end{minipage}
\end{center}
\clearpage


%\thispagestyle{empty}
\begin{center}
\begin{minipage}{\linewidth}
    \centering
%Thesis title
    {\Large APPROVAL PAGE \par}
    %\vspace{1cm}
    {\Large Doctor of Philosophy Dissertation \par}
    \vspace{1cm}
    {\Large \thesisTitle \par}
    \vspace{2cm}
%Author's name
    {\Large Presented by\par}
    {\Large \thesisAuthorName\par}
    \vspace{3cm}
    \begin{tabular}{r l l}
    {\Large Major Advisor} &
    \multicolumn{2}{l}{\noindent\makebox[4in]{\hrulefill}}\\
                           &\makebox[1cm] &{\Large Prof. Sanguthevar
                           Rajasekaran}\\
                           \\
                           \\
    {\Large Associate Advisor} & 
    \multicolumn{2}{l}{\noindent\makebox[4in]{\hrulefill}}\\
                           & &{\Large Prof. Ion M\u{a}ndoiu}\\
                           \\
                           \\
    {\Large Associate Advisor} & 
    \multicolumn{2}{l}{\noindent\makebox[4in]{\hrulefill}}\\
                           & &{\Large Prof. Yufeng Wu}\\
    \end{tabular}
    \vspace{1cm}

    {\Large University of Connecticut\par}
    {\Large 2016\par}
\end{minipage}
\end{center}
\clearpage


\chapter*{Acknowledgments}
This thesis is based on the following publications: \cite{NRPMS14},
\cite{NRPMS15}, \cite{RNSA14}, \cite{NRKM15} and \cite{NRKM16}. I would like to
thank my major advisor, Professor Sanguthevar Rajasekaran, who is also a
co-author on the above papers, for his significant contribution to this thesis
and for the experience of working together for the past few
years.
I would also like to thank my associate advisors Professor Ion M\u{a}ndoiu and
Professor Yufeng Wu for their support and suggestions. I thank also 
Professor Chun-Hsi (Vincent) Huang, Professor Mukul Bansal and Dr. Soumitra
Pal for their participation in the review of this work. I am grateful to
my colleagues from the Applied Algorithms Lab at the Booth Center for Advanced
Technology for the collaborations and sharing of ideas and to
the many excellent teachers in the Computer Science Department at UConn
whose classes I was privileged to attend. 

I would like to acknowledge the
many teachers and colleagues that have influenced me in my undergraduate and
high school years.
In particular, my first steps into computer science were guided by Professor
Doru Popescu Anastasiu from Radu Greceanu National College in Slatina, Romania.
He was also the one who recommended the UConn grad school program to me.
I am grateful for many other influences from professors and former
students training the Romanian National Informatics Olympic team of which I was
a part in my senior high school years. Finally, I would like to thank my family
and friends for their patience and support while I undertook the grad school
program many miles away.

 This work has been supported in part by the following grants:
NSF 0829916, NSF 1447711 and NIH R01-LM010101.

\tableofcontents

\listoffigures
\listoftables
\listofalgorithms

\cleardoublepage\pagenumbering{arabic}

\chapter{Planted Motif Search}
\input{pms.tex}

\chapter{Suffix Array Construction}
\input{sa.tex}

\chapter{Pattern Matching With Mismatches}
\input{kmis.tex}


\cleardoublepage
\phantomsection
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{plain}
\bibliography{pms8-references,pms9-references,sa-references,kmis-references,our-references} 


\end{document}
